---
title: "Collegiate sports"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE, cache=T}
library(flexdashboard)
library(tidyverse)
library(shiny)
library(tidycensus)
library(DT)
library(LogConcDEAD)


if (!dir.exists("data")){
  dir.create("data")
  sports <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-03-29/sports.csv')
  write.csv(sports,"data/sports.csv")
}
sports <- read.csv("data/sports.csv")[,-1]
```

# About

**Purpose**
The purpose of this dashboard is to summary the distribution of collegial sports in United States and analyze correlations between some variables.

**Link to Data**
This data is from TidyTuesday: https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-03-29

Original data is from Equity in Athletics Data Analysis: https://ope.ed.gov/athletics/#/datafile/list

**Quick Guide Video**




# The Data

You can observe all types of sports in US colleges, number of students, and expenditure/revenue for each sport in specific college. 

```{r, echo=FALSE}
DT_sports <- sports %>%
  select(c("year","institution_name","state_cd","ef_male_count",
           "ef_female_count","ef_total_count","sector_name","sum_partic_men","sum_partic_women",
           "rev_men","rev_women","total_rev_menwomen","exp_men","exp_women","total_exp_menwomen","sports")) %>%
  filter(!is.na(total_rev_menwomen),!is.na(total_exp_menwomen),!is.na(sports),!is.na(rev_women),!is.na(exp_women),!is.na(exp_men)) %>%
  rename(state = state_cd, male = ef_male_count,female = ef_female_count,
         total = ef_total_count,man_partic = sum_partic_men,women_partic = sum_partic_women) %>%
  mutate(sports = factor(sports))
```

```{r, echo=FALSE}
DT::renderDataTable({
  DT::datatable(DT_sports,
                caption = htmltools::tags$caption(
                  style = 'caption-side: top; text-align: Left;'))
})
```


# Sports Types (Static)

```{r}
DT_sports %>%
  group_by(sports) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=count, y = reorder(sports, count))) +
  geom_bar(stat='identity') +
  xlab("Count") +
  ylab("Sports") +
  labs(title = "Types of Sports in US Schools", 
       subtitle = "from 2015 to 2019",
       caption = "In descending order of the counts of each sport type")
```

Basketball, Volleyball and Soccer are the top three sports reported in this data which show more than 8000 counts.

# Students Distribution (Static)

```{r}
DT_sports %>%
  group_by(sports) %>%
  summarise(men_prop = sum(man_partic) / (sum(man_partic) + sum(women_partic))) %>%
  ggplot(aes(x=men_prop, y = reorder(sports, men_prop))) +
  geom_bar(stat='identity') +
  xlab("Proportion of Men") +
  ylab("Sports") +
  xlim(0,1)+
  labs(title = "Proportion of Men for each sport", 
       subtitle = "in Collegial Sports in US",
       caption = "In descending order of the proportion of men")
```

Wrestling, Lacrosse, and golf are the top 3 sports with high proportion of men. Equestrian, Volleyball and Rowing are the top 3 sports with high proportion of men. 

# Sports Per States (Interactive)

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
checkboxGroupInput("Selected_states", label = "Choose state",
            choices = c("AL", "AK", "AZ", "AR", "CA", "CO", "CT" ,"DE", "DC", "FL", "GA" ,"HI", "ID","IL" ,"IN" ,"IA", "KS", "KY", "LA" ,"ME", "MD" ,"MA", "MI" ,"MN", "MS", "MO","MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND" ,"OH", "OK", "OR", "PA", "RI", "SC", "SD" ,"TN", "TX", "UT" ,"VT", "VA" ,"WA", "WV" ,"WI" ,"WY" ,"PR", "VI"), selected = c("MD"))

```

Column
-----------------------------------------------------------------------

```{r}
renderPlot({DT_sports %>%
  filter(state == input$Selected_states) %>%
  group_by(sports) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=count, y = reorder(sports, count))) +
  geom_bar(stat='identity') +
  xlab("Count") +
  ylab("Sports") +
  labs(title = "Types of Sports in US Schools", 
       subtitle = "from 2015 to 2019",
       caption = "In descending order of the counts of each sport type")})
```

# Total Expenditure/Revenue (Interactive)

Column {.sidebar}
-----------------------------------------------------------------------

Distribution of Total Expenditure/Revenue for Collegial Sports

```{r}
numericInput("n_breaks", label = "Number of breaks:",5000, min = 10, max = 10000)
numericInput("xlim_end_exp", label = "Maximum value of Expenditure",5e+6, min = 0, max = 28847845)
numericInput("xlim_end_rev", label = "Maximum value of Revenue",5e+6, min = 0, max = 48559421)
```

Column
-----------------------------------------------------------------------

### Distribution of Total Expenditure

```{r}
renderPlot({
  hist(DT_sports$total_exp_menwomen, breaks = as.numeric(input$n_breaks), xlim=c(0,input$xlim_end_exp),
       xlab = "Total Expenditure", main = "Distribution of Total Expenditure")
})

```

### Distribution of Total Revenue

```{r}
renderPlot({
  hist(DT_sports$total_rev_menwomen, breaks = as.numeric(input$n_breaks), xlim=c(0,input$xlim_end_rev),
       xlab = "Total Revenue", main = "Distribution of Total Revenue")
})
```


# Analysis

## Correlation between Expenditure/Revenue

```{r}
top10 <- DT_sports %>%
  group_by(state) %>%
  summarise(count=n())%>%
  top_n(n=8, wt = count) %>%
  select("state")

DT_sports %>% 
  filter(state %in% top10$state) %>%
  ggplot(aes(x=total_exp_menwomen, y = total_rev_menwomen)) +
  geom_line()+
  facet_wrap(~state) +
  xlab("Total Expenditures in USD") +
  ylab("Total Revenue in USD") +
  labs(title = "Relationship between Expenditures and Revenue for sports in US", 
       subtitle = "by states with larger given dataset",
       caption = "There is a positive correlation between the expenditure and revenues")

```

For simpler visualization, we selected top 10 states with larger dataset and it definitely shows positive relationship between expenditures and revenue in sports for students in US schools.

## Distribution of Expenditure in Maryland

```{r}
DT_sports %>% 
  filter(state== "MD", total_exp_menwomen < 1000000) %>%
  ggplot(aes(x=total_exp_menwomen)) +
  geom_histogram() +
  xlab("Expenditure") +
  ylab("Count") +
  labs(title = "Distribution of Expenditure(USD) in Collegial Sports", 
       subtitle = "in Maryland",
       caption = "It's mean is $196,205 and median is $114,522")
```

## MLE estimates for Expenditures/Revenue in Maryland


```{r}
MD <- DT_sports %>% filter(state== "MD", total_exp_menwomen < 200000, total_rev_menwomen < 200000)
MD_df = cbind(MD$total_exp_menwomen,MD$total_rev_menwomen)
lcd <- mlelcd(MD_df)
g <- interplcd(lcd)
plot(lcd, g=g, type="i", uselog=TRUE)
```

